---
title: "final_project_EDA"
author: "Kelly Trinh"
date: "11/29/2020"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE}
library(ggplot2)
library(lattice)
library(corrplot)
library(grid)
library(car)
library(dplyr)
library(carData)
library(GGally)
library(fivethirtyeight)
```

```{r load_data}
# load data
grad <- college_recent_grads
head(grad)
```
```{r variable_names}
# names of all the columns 
names(grad)
```
## 0. Research question 
Question: Given a major category, what’s likely to be my median income as a full-time, year-round worker? 
Model choice: Linear regression. 

## 1. Data cleaning: missing values, data inconsistencies, transformations, outliers

- Variables to exclude: 
Rank: this is by definition, perfectly negatively correlated with median income. 
Major code: this variable doesn't hold any meaningful predictive power. 
Major: we are looking at the major category level, so we will not be looking at major.
p25th: we are predicting median income, so we will not be looking at 25th precentile income. 
p75th: same as above 
Sample size: this variable doesn't hold any meaningful predictive power related to median income. 
Total, men, and women: these three variables are related to the variable "sharewomen". Sharewomen contains information about all these three variables, and since we don't want any highly correlated predictive variables, we will only include sharewoen. 

We will further select more variables when we look at correlations, interactions, and using model selection methods later on. 

```{r understanding_data}
# variables we will consider (so far)
var.include <- c("major_category", "sharewomen", "employed", "employed_fulltime", 
                 "employed_parttime", "employed_fulltime_yearround", "unemployed", "unemployment_rate", 
                 "median", "college_jobs", "non_college_jobs", "low_wage_jobs")
```

- There is one row with missing values for "total", "men", "women", and "sharewomen". We will eliminate this row.

```{r missing_values, echo = FALSE, results = "hide"}
# view row with missing values 
grad[rowSums(is.na(grad)) > 0,]

# remove row with missing value 
grad <- na.omit(grad)
```

- There are some inconsistencies in the data. The number of graduates employed fulltime and employed parttime do not add up to the number of employed graduates. Similarly, the number of college jobs and non-college jobs do not add up to the number of employed graduates. For example, there are some rows where the number of graduates employed fulltime is greater than the number of employed graduates. 

```{r data_inconsistencies, echo = FALSE, results = "hide"}
# view rows where data inconsistencies occur
grad[grad$employed < grad$employed_fulltime, c("employed", "employed_fulltime")]
grad[grad$employed < grad$employed_fulltime_yearround, c("employed", "employed_fulltime_yearround")]
grad[grad$employed < grad$college_jobs + grad$non_college_jobs, c("college_jobs", "non_college_jobs", "employed")]
```

- The target variable is right skewed, so we will log-transform the median income to make it approximately normally distributed. 
```{r transform_data}
# transform categorical data 
grad$major_category <- factor(grad$major_category)

# visualize our target variable: median income 
ggplot(grad, aes(x = median)) + geom_histogram(bins = 100) +
  scale_x_continuous(breaks=seq(min(grad$median), max(grad$median), 10000)) 

# log transform target variable 
grad$log_median = log(grad$median)
# visualize log-transformed variable 
ggplot(grad, aes(x = log_median)) + geom_histogram(bins = 100) +
  scale_x_continuous(breaks=seq(min(grad$log_median), max(grad$log_median), 10000)) 
```
```{r outliers}
# examining outliers
ggplot(data = grad, aes(x = major_category, y = log_median)) + geom_boxplot() + coord_flip()
```
When looking at the outliers, we see that for the major category, the two rows with outliers-"school student counseling" and "library science"-has a very small sample size (only 4 and 2); we will leave these two rows in the dataset for now, and will choose whether to eliminate them later when we perform diagnostics to access whether these rows signficantly affect the outputs of the OLS regression. 

```{r outliers_ct, echo = FALSE, results = "hide"}

# look at the outliers in detail

# engineering: petroleum engineering
grad.eng <- grad[grad$major_category == "Engineering", ]
grad.eng[grad.eng$log_median == max(grad.eng$log_median), ]

# biology and life sciences: zoology, high proportion of non-college jobs 
grad.bio <- grad[grad$major_category == "Biology & Life Science", ]
grad.bio[grad.bio$log_median == min(grad.bio$log_median), ]

# arts
grad.arts <- grad[grad$major_category == "Arts", ]
grad.arts[grad.arts$log_median == max(grad.arts$log_median), ]

# communications & journalism 
grad.com <- grad[grad$major_category == "Communications & Journalism", ]
grad.com[grad.com$log_median == min(grad.com$log_median), ]

# education
grad.edu <- grad[grad$major_category == "Education" & order(grad$log_median, decreasing = TRUE), ]
grad.edu[c(1, nrow(grad.edu) - 1, nrow(grad.edu) ), ]
```
## 2. Examine correlation between variables 

Some interesting observations include:
- Log_median has strong negative correlations with "sharewomen".
- Variables related to job type (for example, employed_fulltime, non_college_jobs, etc.) are highly correlated. Such high correlation can lead to multicollinearity problems in the data. We therefore decided to transform data. We created the variables "fulltime_yearround_prop", "low_wage_prop", and "college_prop".These variables would have weaker correlations with one another compared to the original variables "employed_fulltime_yearround", "low_wage_jobs", and "college_jobs". 

```{r correlation}
# data with only continuous variables that will be included in the model 
# (var.include is defined above in the data cleaning section)
cont <- grad[, c("log_median", var.include[2:length(var.include)])]

# visualizing the correlation matrix (var.include )
corrplot(cor(cont), order = "AOE", type = "upper")
```
- After transforming the data, we observe a few interesting things: 
"Low_wage_prop" has strong correlations with the response variable "log_median", and weaker correlations with other predictor variables when compared to "low_wage_jobs". The same occurs when variable "college_prop" and "fulltime_yearround_prop". However, some multicollinearity still exists. 

```{r transform_data_correlation}
# transform employed_fulltime_yearround
grad$fulltime_yearround_prop <- grad$employed_fulltime_yearround / grad$employed

# transform employed_parttime
grad$parttime_prop <- grad$employed_parttime / grad$employed

# transform college_jobs
grad$college_prop <- grad$college_jobs / (grad$college_jobs + grad$non_college_jobs)

# transform non_college_jobs
grad$noncollege_prop <- grad$non_college_jobs / (grad$college_jobs + grad$non_college_jobs)

# transform low_wage_jobs
grad$low_wage_prop <- grad$low_wage_jobs / grad$employed

# re-examining the correlation matrix after transforming the data
trans <- grad[, c("log_median", "fulltime_yearround_prop", "college_prop", "low_wage_prop", "parttime_prop", "noncollege_prop", var.include[2:length(var.include)])]

trans <- na.omit(trans)
corrplot(cor(trans), order = "AOE", type = "upper")


# omit na values 
grad <- na.omit(grad)
# examine the correlation numbers of the newly created variables
cor(trans[, c("log_median", "fulltime_yearround_prop", "college_prop", "low_wage_prop", "parttime_prop", "noncollege_prop")])

```
After looking at multicollinearity, we have further refined the list of variables that we will consider during our model selection process. 

- We will not include "employed", "employed_fulltime", "employed_parttime", "employed_fulltime_yearround" because these are highly correlated, so we will include "fulltime_yearround_prop" as this variable has low correlation with other response variables and comes from a variable that is highly correlated with the variables "employed", "employed_fulltime", and "employed_parttime". This means that the included variable would capture the information of all the four variables that will be excluded. 
- We will not include "unemployed" because this variable causes multicollinearity, and its information is captured by "unemployment rate". 
- We will not include "college_jobs", "non_college_jobs", and "low_wage_jobs, as these are highly correlated with one another, and their information are captured by variables "college_prop" and "low_wage_prop". 
- Among the transformed variables, we will not include "parttime_prop" and "noncollege_prop" as these variables are by definition highly correlated with variables "fulltime_yearround_prop" and "college_prop", respectively. Also, they are less relevant to our analysis as we are trying to predict the median income of fulltime, yearround workers. 


```{r select_variables, echo = FALSE, results = "hide" }

# selected variables before looking at correlations and transformations
var.include

# selected variables after looking at correlations 
var.after_cor <- c("major_category", "sharewomen", "unemployment_rate", "log_median", "fulltime_yearround_prop", "college_prop", "low_wage_prop")
```

##3. Interactions

The variable "sharewomen" and "fulltime_yearround_prop" seem to have some interesting interactions with the variable "major_category". 

```{r coplots, echo = FALSE, warning = "hide"}


# these co-plots do not contain some major categories because there are two few rows for each of those major categories

# share women and median income 
xyplot(log_median ~ sharewomen | major_category, data = grad[!(grad$major_category %in% c("Communications & Journalism", "Interdisciplinary", "Law & Public Policy")), ],
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# unemployment and median income 
xyplot(log_median ~ unemployment_rate | major_category, data = grad[!(grad$major_category %in% c("Communications & Journalism", "Interdisciplinary", "Law & Public Policy")), ],
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# full time year round and median income 
xyplot(log_median ~ fulltime_yearround_prop | major_category, data = grad[!(grad$major_category %in% c("Communications & Journalism", "Interdisciplinary", "Law & Public Policy")), ],
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# low wage and median income
xyplot(log_median ~ low_wage_prop | major_category, data = grad[!(grad$major_category %in% c("Communications & Journalism", "Interdisciplinary", "Law & Public Policy")), ],
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# college jobs and median income
xyplot(log_median ~ college_prop | major_category, data = grad[!(grad$major_category %in% c("Communications & Journalism", "Interdisciplinary", "Law & Public Policy")), ],
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})
```
