---
title: "Stat 151a Project"
author: "Kelly Trinh, Hanfei Sun, and Nathan Lam"
output: pdf_document
---

https://www.rdocumentation.org/packages/fivethirtyeight/versions/0.6.1/topics/college_recent_grads


```{r Libraries, message=FALSE, warning=FALSE}
library(fivethirtyeight) #source of data
library(corrplot) #correlation plot
library(dplyr) #data frame manipulation
library(ggplot2) #plotting
library(leaps)#variable selection: adjr2, bic, cp
library(olsrr)#variable selection: aic, p-val
library(fastDummies)
library(faraway)
library(caret)
```

```{r}
crg <- college_recent_grads
cont <- select_if(crg, is.numeric)
cont["college_job_prop"] <- cont$college_jobs / (cont$college_jobs + cont$non_college_jobs)
cont["full_time_yearround_prop"] <- cont$employed_fulltime_yearround / cont$employed
cont <- na.omit(cont)
corrplot(cor(cont), type = "upper", order = "AOE")
crg["log_median"] <- log(crg$median)
crg["college_job_prop"] <- crg$college_jobs / (crg$college_jobs + crg$non_college_jobs)
crg["full_time_yearround_prop"] <- crg$employed_fulltime_yearround / crg$employed
crg <- na.omit(crg)

# temporary data frame to use later 
temp <- crg 


crg <- dummy_cols(crg, select_columns = "major_category", remove_selected_col = TRUE,remove_first_dummy = TRUE)
column_names <- make.names(names(crg),unique=TRUE)
colnames(crg) <- column_names

```

```{r}
ggplot(aes(x = major_category, y = log_median), data = temp) + geom_boxplot() + coord_flip()
```

```{r Variable Selection}
Y <- as.matrix(crg$log_median)

X <- as.matrix(subset(crg,select=-c(log_median,median, major_code, p25th, p75th, major)))
# Check for linear dependencies, remove "men" or "total"
alias(lm(Y~X))
X <- as.matrix(subset(crg,select=-c(log_median,median, major_code, p25th, p75th, major, men)))
```

```{r}
#Forward Selection
model <- regsubsets(x = X, y = Y, method = "forward", nbest = 1) %>% summary()

adjustr2 <- model$adjr2
BIC <- model$bic
mallow_Cp <- model$cp


cbind(model$which[which.max(adjustr2),], model$which[which.min(BIC),],model$which[which.min(mallow_Cp),])
```

```{r}
###################
#Backward Selection

model <- regsubsets(x = X, y = Y, method = "backward", nbest = 1) %>% summary()


adjustr2 <- model$adjr2
BIC <- model$bic
mallow_Cp <- model$cp
cbind(model$which[which.max(adjustr2),], model$which[which.min(BIC),],model$which[which.min(mallow_Cp),])
```

### Cross Validation

```{r}
# Backward Adjusted R-squared CV
tc <- trainControl(method = "LOOCV")
b_adjr2_mod_cv <- train(log_median ~ employed + employed_fulltime_yearround + 
                          college_job_prop +
                          major_category_Biology...Life.Science+
                          major_category_Computers...Mathematics + 
                          rank + sharewomen + major_category_Engineering , 
                        data = training_data, method = "lm", trControl = tc)
b_adjr2_mod_cv.rmse <- b_adjr2_mod_cv$results[,"RMSE"]

```


```{r}
# Backward BIC CV
b_bic_mod_cv <- train(log_median ~ rank + college_job_prop +  major_category_Engineering, data = training_data, method = "lm", trControl = tc )
b_bic_mod_cv.rmse <- b_bic_mod_cv$results[, "RMSE"]
```

```{r}
# Backward Mallow's Cp CV
b_cp_mod_cv <- train(log_median ~ employed + employed_fulltime_yearround + 
                       college_job_prop +  major_category_Biology...Life.Science
                     + major_category_Computers...Mathematics + rank +
                       major_category_Engineering + sharewomen , data = training_data , 
                     method = "lm", trControl = tc)
b_cp_mod_cv.rmse <- b_cp_mod_cv$results[, "RMSE"]
```


```{r}
# Forward Adjusted R-squared CV
f_adjr2_model_cv <- train(log_median ~ rank + unemployment_rate + 
                           college_job_prop + major_category_Arts + 
                           major_category_Biology...Life.Science  + 
                           major_category_Engineering  + 
                           major_category_Humanities...Liberal.Arts  + 
                           major_category_Industrial.Arts...Consumer.Services , 
                           data = training_data, method = "lm", trControl = tc)

f_adjr2_model_cv.rmse <- f_adjr2_model_cv$results[, "RMSE"]

# Forward BIC CV
f_BIC_model_cv <- train(log_median ~ rank + college_job_prop + 
                           major_category_Engineering, 
                       data = training_data, method = "lm", trControl = tc)

f_BIC_model_cv.rmse <- f_BIC_model_cv$results[, "RMSE"]

# Forward Mallow's Cp CV
f_Cp_model_cv <- train(log_median ~ rank + unemployment_rate + 
                          college_job_prop + major_category_Arts + 
                         major_category_Engineering + 
                         major_category_Humanities...Liberal.Arts, 
                      data = training_data, method = "lm", trControl = tc)

f_Cp_model_cv.rmse <- f_Cp_model_cv$results[, "RMSE"]
```

Summary table description: 

`forward.rmse` and `backward.rmse` are columns with the RMSE from prediction on the testing dataset depending on either forward selection or backward selection. 
Each of the row represent the criteria in which to choose the model from either forward or backward selection: `Adjusted R squared`, `BIC`, `Mallow's Cp`. 

```{r CV_rmse_summary}

criterias <- c("Adjusted R squared", "BIC", "Mallow's Cp")
forward.rmse <- c(f_adjr2_model_cv.rmse, f_BIC_model_cv.rmse, f_Cp_model_cv.rmse)
backward.rmse <- c(b_adjr2_mod_cv.rmse, b_bic_mod_cv.rmse, b_cp_mod_cv.rmse)

# table of RMSE of each model 
data.frame(criterias, forward.rmse, backward.rmse)
```
```{r}
#interactions

# these co-plots do not contain some major categories because there are two few rows for each of those major categories

# temp dataframe for co-plots
temp.training <- temp[temp$rank %in% training_data$rank,]
temp.coplot <- temp.training[temp.training$major_category %in% 
                               c("Arts", 
                                 "Biology & Life Science",
                                 "Computers & Mathematics", 
                                 "Engineering"), ]


# share women and median income 
xyplot(log_median ~ sharewomen | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# rank and median income 
xyplot(log_median ~ rank | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# college job proportion and median income 
xyplot(log_median ~ college_job_prop | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# employed fulltime and median income 
xyplot(log_median ~ employed_fulltime_yearround | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})


```
```{r interactions}

# full model with all the interaction terms 
model.interacts <- lm(log_median ~ 
                        rank + sharewomen + 
                        employed + employed_fulltime_yearround + 
                        college_job_prop +
                        major_category_Biology...Life.Science+
                        major_category_Computers...Mathematics + 
                        major_category_Engineering +
                        major_category_Biology...Life.Science * rank+
                        major_category_Computers...Mathematics * rank+ 
                        major_category_Engineering * rank +
                        major_category_Biology...Life.Science * sharewomen+
                        major_category_Computers...Mathematics * sharewomen + 
                        major_category_Engineering * sharewomen +
                        major_category_Biology...Life.Science * employed_fulltime_yearround+
                        major_category_Computers...Mathematics * employed_fulltime_yearround+ 
                        major_category_Engineering * employed_fulltime_yearround+
                        major_category_Biology...Life.Science * college_job_prop+
                        major_category_Computers...Mathematics *college_job_prop + 
                        major_category_Engineering *college_job_prop, 
                        data = training_data)

# anova test
Anova(model.interacts)

# reduced model after running Anova test 
model.inter_reduced <- lm(log_median ~ 
                        rank + sharewomen + 
                        employed + employed_fulltime_yearround + 
                        college_job_prop +
                        major_category_Biology...Life.Science+
                        major_category_Computers...Mathematics + 
                        major_category_Engineering +
                        major_category_Engineering * rank +
                        major_category_Engineering * sharewomen,  
                        data = training_data)
   
   
anova(model.interacts, model.inter_reduced)


```

```{r compare_models}
# compare model with and without interactions 
interactions_cv <- train(log_median ~ 
                        rank + sharewomen + 
                        employed + employed_fulltime_yearround + 
                        college_job_prop +
                        major_category_Biology...Life.Science+
                        major_category_Computers...Mathematics + 
                        major_category_Engineering +
                        major_category_Engineering * rank +
                        major_category_Engineering * sharewomen, 
                      data = training_data, method = "lm", trControl = tc)

interactions_cv.rmse <- interactions_cv$results[, "RMSE"]

interactions_cv.rmse
```

### Diagnostics 

```{r}
plot(model.inter_reduced)
```
```{r}
# outlier rows 
training_data[c(14, 33, 74, 87),]

# refit model with the outliers 
interactions_no_outliers <- train(log_median ~ 
                        rank + sharewomen + 
                        employed + employed_fulltime_yearround + 
                        college_job_prop +
                        major_category_Biology...Life.Science+
                        major_category_Computers...Mathematics + 
                        major_category_Engineering +
                        major_category_Engineering * rank +
                        major_category_Engineering * sharewomen, 
                      data = training_data[-c(14, 33, 74, 87),], method = "lm", trControl = tc)


# see CV RMSE
interactions_no_outliers.rmse <- interactions_no_outliers$results[, "RMSE"]
interactions_no_outliers.rmse


plot(lm(log_median ~ 
                        rank + sharewomen + 
                        employed + employed_fulltime_yearround + 
                        college_job_prop +
                        major_category_Biology...Life.Science+
                        major_category_Computers...Mathematics + 
                        major_category_Engineering +
                        major_category_Engineering * rank +
                        major_category_Engineering * sharewomen, 
                      data = training_data[-c(14, 33, 74, 87),]))

```
