---
title: "stat151a"
author: "Kelly Trinh"
date: "12/9/2020"
output: html_document
---

### Research question 

Question: Given a major category, what’s likely to be my median income as a full-time, year-round worker? 
Model choice: Linear regression. 

### Methodology 

1) EDA 
- NaN values.
- Correlation.
- Transform variables.

2) Variable selection without interactions 
- Standardize continuous coefficients.
- Train-test split.
- Forward/backward selection based on Mallow's Cp, adjusted R squared, and 
BIC output a few best models. 
- Choose the model with the smallest RMSE based on predictions on test set.

3) Variable selection with interactions 
- Interaction plots (using training dataset).
- Take model selected from step 2 and add interaction terms; run Anova. 
- Compare model with interaction terms after Anova and without interactions using 
RMSE based on predictions on test set. 

4) Diagnostics 
- Diagnostics on outliers and refit model after eliminating outliers. 
- Diagnostics on assumptions. 

5) Inference 
- Bootstrapping coefficients to comment on stability of coefficients.
- Interpret the coefficients and report confidence interval. 

6) Predictions
- Construct prediction inteval using a row or two from the testing dataset. 

```{r Libraries, message=FALSE, warning=FALSE}
library(fivethirtyeight) #source of data
library(corrplot) #correlation plot
library(dplyr) #data frame manipulation
library(ggplot2) #plotting
library(leaps)#variable selection: adjr2, bic, cp
library(olsrr)#variable selection: aic, p-val
library(car) #used for diagnostics
library(fastDummies)
library(faraway)
library(caret)
```

```{r data_cleaning}
# load data 
crg <- college_recent_grads

# transform variables 
crg["log_median"] <- log(crg$median)
crg["college_job_prop"] <- crg$college_jobs / (crg$college_jobs + crg$non_college_jobs)
crg["full_time_yearround_prop"] <- crg$employed_fulltime_yearround / crg$employed

# omit NaN values 
crg <- na.omit(crg)

# temporary data frame 
temp <- crg

# transform categorical variables
crg <- dummy_cols(crg, select_columns = "major_category", remove_selected_col = TRUE,remove_first_dummy = TRUE)
column_names <- make.names(names(crg),unique=TRUE)
colnames(crg) <- column_names
```

### EDA
```{r correlation}
# continuous variables 
cont <- select_if(crg, is.numeric)
corrplot(cor(cont), type = "upper", order = "AOE")
```

```{r visualizations}
# plot response variable 
ggplot(grad, aes(x = median)) + geom_histogram(bins = 100) +
  scale_x_continuous(breaks=seq(min(grad$median), max(grad$median), 10000)) 

# plot response variable against major categories 
ggplot(aes(x = major_category, y = log(median)), data = college_recent_grads) + geom_boxplot() + coord_flip()

```

### Variable selection without interactions

```{r train_test_split}



set.seed(11)

#randomly take 2:8 of data for training
training_size <- sample(dim(crg)[1],nrow(crg)*0.8)
training_data <- crg[training_size,]
testing_data <- crg[-training_size,]

# reserve three rows from testing set to construct prediction interval 
predict.rows <- sample_n(testing_data, 3) 
```

```{r prepare_data}


Y <- as.matrix(training_data$log_median)

X <- as.matrix(subset(training_data,select=-c(log_median,median, major_code, p25th, p75th, major)))

# Check for linear dependencies, remove "men" or "total"
alias(lm(Y~X))
X <- as.matrix(subset(training_data,select=-c(log_median,median, major_code, p25th, p75th, major, men)))
``` 

```{r forward}
model <- regsubsets(x = X, y = Y, method = "forward", nbest = 1) %>% summary()

adjustr2 <- model$adjr2
BIC <- model$bic
mallow_Cp <- model$cp

cbind(model$which[which.max(adjustr2),], model$which[which.min(BIC),],model$which[which.min(mallow_Cp),])

```

```{r backward}
model <- regsubsets(x = X, y = Y, method = "backward", nbest = 1) %>% summary()

adjustr2 <- model$adjr2
BIC <- model$bic
mallow_Cp <- model$cp
cbind(model$which[which.max(adjustr2),], model$which[which.min(BIC),],model$which[which.min(mallow_Cp),])

```

```{r CV_backward}
# Backward Adjusted R-squared CV
tc <- trainControl(method = "LOOCV")
b_adjr2_mod_cv <- train(log_median ~ employed + employed_fulltime_yearround + 
                          college_job_prop +
                          major_category_Biology...Life.Science+
                          major_category_Computers...Mathematics + 
                          rank + sharewomen + major_category_Engineering , data = crg, method = "lm",
               trControl = tc)
b_adjr2_mod_cv.rmse <- b_adjr2_mod_cv$results[,"RMSE"]


# Backward BIC CV
b_bic_mod_cv <- train(log_median ~ rank + college_job_prop +  major_category_Engineering, data = crg, method = "lm", trControl = tc )
b_bic_mod_cv.rmse <- b_bic_mod_cv$results[, "RMSE"]

# Backward Mallow's Cp CV
b_cp_mod_cv <- train(log_median ~ employed + employed_fulltime_yearround + 
                       college_job_prop +  major_category_Biology...Life.Science
                     + major_category_Computers...Mathematics + rank +
                       major_category_Engineering + sharewomen , data = crg , 
                     method = "lm", trControl = tc)
b_cp_mod_cv.rmse <- b_cp_mod_cv$results[, "RMSE"]

```

```{r CV_forward}
# Forward Adjusted R-squared CV
f_adjr2_model_cv <- train(log_median ~ rank + unemployment_rate + 
                           college_job_prop + major_category_Arts + 
                           major_category_Biology...Life.Science  + 
                           major_category_Engineering  + 
                           major_category_Humanities...Liberal.Arts  + 
                           major_category_Industrial.Arts...Consumer.Services , 
                           data = crg, method = "lm", trControl = tc)

f_adjr2_model_cv.rmse <- f_adjr2_model_cv$results[, "RMSE"]

# Forward BIC CV
f_BIC_model_cv <- train(log_median ~ rank + college_job_prop + 
                           major_category_Engineering, 
                       data = crg, method = "lm", trControl = tc)

f_BIC_model_cv.rmse <- f_BIC_model_cv$results[, "RMSE"]

# Forward Mallow's Cp CV
f_Cp_model_cv <- train(log_median ~ rank + unemployment_rate + 
                          college_job_prop + major_category_Arts + 
                         major_category_Engineering + 
                         major_category_Humanities...Liberal.Arts, 
                      data = crg, method = "lm", trControl = tc)

f_Cp_model_cv.rmse <- f_Cp_model_cv$results[, "RMSE"]

```

Summary table description: 

`forward.rmse` and `backward.rmse` are columns with the RMSE from prediction on the testing dataset depending on either forward selection or backward selection. 
Each of the row represent the criteria in which to choose the model from either forward or backward selection: `Adjusted R squared`, `BIC`, `Mallow's Cp`. 

```{r CV_rmse_summary}

criterias <- c("Adjusted R squared", "BIC", "Mallow's Cp")
forward.rmse <- c(f_adjr2_model_cv.rmse, f_BIC_model_cv.rmse, f_Cp_model_cv.rmse)
backward.rmse <- c(b_adjr2_mod_cv.rmse, b_bic_mod_cv.rmse, b_cp_mod_cv.rmse)

# table of RMSE of each model 
data.frame(criterias, forward.rmse, backward.rmse)

```
### Variable selection with interactions 

The 2 models with the lowest RMSE: 
selected with backward selection based on adjusted R squared and Mallow's Cp. 


```{r selected_models_backward_forward}

# selected models from 

log_median ~ employed + employed_fulltime_yearround + 
                          college_job_prop +
                          major_category_Biology...Life.Science+
                          major_category_Computers...Mathematics + 
                          rank + sharewomen + major_category_Engineering 


log_median ~ employed + employed_fulltime_yearround + 
                       college_job_prop +  major_category_Biology...Life.Science
                     + major_category_Computers...Mathematics + rank +
                       major_category_Engineering + sharewomen 

```
```{r interaction_plots}
# these co-plots do not contain some major categories because there are two few rows for each of those major categories

# temp dataframe for co-plots
temp.training <- temp[temp$rank %in% training_data$rank,]
temp.coplot <- temp.training[temp.training$major_category %in% 
                               c("Biology & Life Science", 
                                 "Computers & Mathematics", 
                                 "Engineering"), ]

# share women and median income 
xyplot(log_median ~ sharewomen | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# unemployment and median income 
xyplot(log_median ~ unemployment_rate | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# full time year round and median income 
xyplot(log_median ~ fulltime_yearround_prop | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# low wage and median income
xyplot(log_median ~ low_wage_prop | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

# college jobs and median income
xyplot(log_median ~ college_prop | major_category, data = temp.coplot ,
       panel = function(x, y) {
panel.xyplot(x, y, type = c("p", "smooth", "r"))
})

```